```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE, 
  message=FALSE, 
  warning=FALSE, 
  fig.width = 10,
  fig.align = "center"
  )

library( here )      # for the directory
library( ggraph )    # for plotting
library( igraph )    # for working with graphs
library( ggplot2 )   # for plotting
library( gridExtra ) # for plotting multiple plots
library( grid )      # to include the null graphical object
library( reshape2 )  # for reshaping the matrix data for a plot

```

# Closeness Centrality

In the prior lecture, we examined the idea of "centrality" for nodes and "centralization" for graphs, focusing attention on the conceptual aspects of what it means to be *central* as well as the ways in which we can operationalize the concept using *degree centrality*. In this chapter we want to continue thinking about what it means to be central, but through a different conceptual lens: **closeness**. As a crime analyst, examining closeness centrality can provide crucial insights into the efficiency and influence of individuals within a group that may go beyond what you can glean from examining degree centrality. For instance, @bright2015illuminating study of drug trafficking networks illustrates how closeness centrality played a pivotal role in understanding the structure and function of illicit organizations. In this study, @bright2015illuminating examined an Australian drug trafficking network and found that individuals with higher closeness centrality were not necessarily the most visible or directly connected (in terms of degree centrality), but held critical positions because of their proximity to others in the network. These individuals could quickly access and disseminate information across the group, making them key actors in maintaining the flow of drugs, money, and communication. By identifying those with high closeness centrality, law enforcement was able to target individuals who had the potential to disrupt the network's operations more effectively than simply targeting those with the most connections.

By the end of this chapter, you should be able to:

-   Explain the conceptualization of "centrality" as it pertains to *closeness centrality*.
-   Calculate closeness centrality scores for a set of nodes.
-   Calculate a closeness centralization score for a graph.

```{r, graphs-define}

# ----
# This creates the graphs for the plots below

# Create the graph objects
ugraph  <- graph_from_data_frame( 
  data.frame( 
    from = c( 1, 2, 3, 3, 4 ),
    to   = c( 2, 3, 4, 5, 5 ) ), 
  directed = FALSE )

digraph  <- graph_from_data_frame( 
  data.frame( 
    from = c( 1, 2, 3, 3, 4, 4, 5, 5  ),
    to   = c( 2, 3, 4, 5, 3, 5, 3, 4  ) ), 
  directed = TRUE )

star_net <- graph_from_data_frame(
  data.frame(
    from = c( 1, 1, 1, 1 ),
    to   = c( 2, 3, 4, 5 ) ),
  directed = FALSE )

circle_net <- graph_from_data_frame(
  data.frame(
    from = c( 1, 2, 3, 4, 5 ),
    to   = c( 2, 3, 4, 5, 1 ) ),
  directed = FALSE )

# Set the random seed to render the same plot
set.seed( 507 )

# Set a fixed layout using the Fruchterman-Reingold layout
layout <- layout_with_fr( ugraph )
dilayout <- layout_with_fr( digraph )

starlayout <- layout_with_fr( star_net )
circlelayout <- layout_with_fr( circle_net )


# Set the labels
custom_labels <- c( "Jen","Tom","Bob","Leaf","Jim" )

# Assign the labels to the graph nodes
V( ugraph )$name <- custom_labels
V( digraph )$name <- custom_labels

```

```{r, plots-define}

# create the undirected plots
u_graph <-
  ggraph( ugraph, 
          layout = layout ) +                  
  geom_edge_link( color = "black", width = 0.8 ) +  
  geom_node_point( color = "#28a88d", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5 ) +
  ggtitle( "Undirected Graph" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
theme_void() 

star_graph <-
  ggraph( star_net, 
          layout = starlayout ) +                  
  geom_edge_link( ) +  
  geom_node_point( color = "#28a88d", size = 10 ) +
  ggtitle( "Graph w/ High Degree Centralization" ) + 
theme_void() 

circle_graph <-
  ggraph( circle_net, 
          layout = circlelayout ) +                  
  geom_edge_link( ) +  
  geom_node_point( color = "#28a88d", size = 10 ) +
  ggtitle( "Graph w/ Low Degree Centralization" ) + 
theme_void() 


# create the directed graphs
d_graph <- 
  ggraph( digraph, 
          layout = dilayout ) +                  
  geom_node_point( color = "#fc23fc", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5) +
  geom_edge_link( aes( start_cap = label_rect( node1.name ), 
                     end_cap = circle( 5, 'mm' ) ),
                 arrow = arrow( length = unit( 0.02, "npc" ) ), 
                 color = "black", width = 0.8) +
  ggtitle( "Directed Graph" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  theme_void()  

```

## Case Study: Terror Networks and Closeness

@sageman2004understanding studied global Islamist terrorist networks and demonstrated how individuals with high closeness centrality played a key role in maintaining the cohesion and communication efficiency within these decentralized groups. These key players, even though they may not have had the most direct connections (high degree centrality), were crucial in maintaining the operational efficiency of the network. By targeting those with high closeness centrality, counter-terrorism efforts were able to more effectively disrupt communication channels and reduce the operational capabilities of the group.

### Closeness Centrality Terms

How *close* is a node to other nodes? *Closeness* centrality measures how near a node is to the other nodes in the network. That is, how "close" the nodes are. What we mean by "close" will differ based on whether the graph is undirected or directed. By the conceptual definition is the same. How close nodes are in a graph is based on the inverse of the distance of each node to every other node.

Terminology:

-   A **path** is an edge or sequence of edges that connect two nodes.
-   A **geodesic** is the shortest path between two nodes.
-   The **distance**, $d(n_i,n_j)$, is the length of the path between *i* and *j*.

Another way of saying this is that if we now how "far apart" two nodes are (i.e. the distance), then we can take the inverse of that and know how "close" the nodes are to each other.

## Undirected Graphs

### Closeness Centrality

For an undirected binary graph, the closeness is:

$$C_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1}$$

This reads as: Find the distance between two nodes *i* and *j*, $d(n_i,n_j)$, sum them, and then take the inverse.

Another way to write this is as:

$$C_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} = \frac{1}{[\sum\limits_{j=1}^g d(n_i,n_j)]}$$

In words, the closeness is 1 divided by the sum of the distances (called the *reciprocal*).

To find the closeness centrality score, we need to first find the distances between each node. We can express these distances using a matrix, just like we did with the sociomatrix before.

Let's take a look at this with some data. Recall our example network with Jen, Tom, Bob, Leaf, and Jim:

```{r, adj-undirected-define}

# create an adjacency matrix from the graph
adj_matrix <- as_adjacency_matrix( ugraph, sparse = FALSE )

# melt the matrix into a long format suitable for ggplot
df <- melt( adj_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
df$Var2 <- factor( df$Var2, levels = rev( rownames( adj_matrix ) ) )

# plot the adjacency matrix
adj_to_plot <- ggplot( df, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  geom_text( aes( label = value ), color = "black" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Adjacency Matrix for Undirected Graph" )

```

```{r}
ugraph
```

How far is Jen from Tom? How far is Jen from Bob? In other words, how many edges do we count in going from Jen to Tom? From Jen to Bob?

```{r}

# create the distance matrix
d_matrix <- distances( ugraph )

# melt the matrix into a long format suitable for ggplot
dfd <- melt( d_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
dfd$Var2 <- factor( df$Var2, levels = rev( rownames( adj_matrix ) ) )

# plot the blank distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

For our graph, we want to first build a *distance matrix* which shows the value of the geodesic (i.e. shortest path) from *i* to *j*. We can do this by looking at each node and seeing the fewest steps that are needed to reach another node. We can see that Jen can reach Tom in 1 step (i.e. $L_{Jen,Tom}$) and that Jen can reach Bob in 2 steps (i.e. $L_{Jen,Tom}$ & $L_{Tom,Bob}$).

```{r}

# dummy the distances for Jen
dfd$Jen <- ifelse( rownames( dfd ) %in% c( 1:5 ), "show", "dont" )

# plot the distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#28a88d", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

Since the graph is undirected, the distance matrix is symmetric about the diagonal (just like we saw with an adjacency matrix). That is, $d(n_i,n_j) = d(n_j,n_i)$. This gives us:

```{r}

# dummy the distances for Jen
dfd$Jen <- ifelse( rownames( dfd ) %in% c( 1:5, 6, 11, 16, 21 ), "show", "dont" )

# plot the distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#28a88d", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

Now, what about the rest of the distance matrix? If we follow what we did above for everyone else, then we get the following distance matrix:

```{r}

# plot the distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  geom_text( aes( label = value ), color = "black" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

Now, it we take the sum of the rows, this gives us $\sum\limits_{j=1}^g d(n_i,n_j)$. Note that since the graph is undirected, the column sum would give us the same values. Since we have $\sum\limits_{j=1}^g d(n_i,n_j)$, all we need to do now is take the reciprocal by dividing 1 by $\sum\limits_{j=1}^g d(n_i,n_j)$:

$$C_C(n_i) =  \frac{1}{[\sum\limits_{j=1}^g d(n_i,n_j)]}$$

That gives us the closeness centrality score for each individual:

| Node | Distance |  Closeness  |
|:----:|:--------:|:-----------:|
| Jen  |    9     | 1/9 = 0.111 |
| Tom  |    6     | 1/6 = 0.167 |
| Bob  |    5     | 1/5 = 0.200 |
| Leaf |    7     | 1/7 = 0.143 |
| Jim  |    7     | 1/7 = 0.143 |

### Standardization

As discussed in the prior chapter, centrality measures are sensitive to the size of the graph, *g*. In the case of *closeness* centrality, having to sum over more nodes will push scores closer to zero. Think about it: if we added more rows and columns to our distance matrix, that will increase the denominator, making the quotient smaller. This means that we can't compare closeness centrality scores across graphs of different sizes. *Solution*?

Standardize! If we want to compare nodes in different sized networks, then we just take into account the number of nodes and the maximum possible nodes to which *i* could be connected. In an undirected graph, this is *g-1*.

Thus, we can calculate a standardized closeness centrality for an undirected graph as:

$$C'_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} \times [g-1]$$

For this graph, the denominator, *g-1* is $5 - 1 = 4$. Plugging that in we get: $C'_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} \times 4$.

| Node | Distance |  Closeness  | Standardized Closeness |
|:----:|:--------:|:-----------:|:----------------------:|
| Jen  |    9     | 1/9 = 0.111 |    0.111\*4 = 0.444    |
| Tom  |    6     | 1/6 = 0.167 |    0.167\*4 = 0.668    |
| Bob  |    5     | 1/5 = 0.200 |    0.200\*4 = 0.800    |
| Leaf |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |
| Jim  |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |

Standardization is particularly important with *closeness* centrality because of the sensitivity of the measure to the size of the network. Take for example, the following networks:

```{r}

# set up the objects for the plot
graph_a <- graph_from_adjacency_matrix( 
  matrix(
    c( 0, 1, 0,
      1, 0, 1,
      0, 1, 0),
    byrow = TRUE, ncol = 3 ), 
  mode = "undirected" 
  )

graph_b <- graph_from_adjacency_matrix( 
  matrix(
    c( 0, 1, 1, 1,
      1, 0, 1, 1,
      1, 1, 0, 1,
      1, 1, 1, 0 ),
    byrow = TRUE, ncol = 4 ), 
  mode = "undirected" 
)

# set the vertex and edge properties for Network A
V( graph_a )$color <- c( "lightblue", "grey80", "grey80" )
V( graph_a )$size  <- 10
V( graph_a )$label <- c( "0.33", "", "" )
V( graph_a )$label_s <- c( "0.66","","" )
V( graph_a )$label_s2 <- c( "0.66","1","0.66" )
E( graph_a )$color <- "grey50"

# set the vertex and edge properties for Network B
V( graph_b )$color <- c( "orange", "grey80", "grey80", "grey80" )
V( graph_b )$size <- 10
V( graph_b )$label <- c( "0.33", "", "", "" )
V( graph_b )$label_s <- c( "1", "", "", "" )
V( graph_b )$label_s2 <- c( "1", "1", "1", "1" )
E( graph_b )$color <- "grey50"

# set the seed for the layout
set.seed( 507 ) 

# set the layouts
layout_a <- layout_with_fr( graph_a )
layout_b <- layout_with_fr( graph_b )

```

```{r}

# Plot Network A using ggraph
p1 <- ggraph( graph_a, layout = layout_a ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_a )$color ), size = I( V( graph_a )$size ) ) ) +
  geom_node_text( aes( label = V( graph_a )$label ), size = 5 ) +
  labs( title = "Network A: Raw Closeness Centrality" ) +
  theme_void()

# Plot Network b using ggraph
p2 <- ggraph( graph_b, layout = layout_b ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_b )$color ), size = I( V( graph_b )$size ) ) ) +
  geom_node_text( aes( label = V( graph_b )$label ), size = 5 ) +
  labs( title = "Network B: Raw Closeness Centrality" ) +
  theme_void()

grid.arrange( p1, nullGrob(), p2, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

The plot above shows two different graphs with the *raw* closeness centrality scores for two nodes. Note that they have the exact same score. Yet, would you say they are equally close to other nodes in the network? In other words, that their closeness to other nodes is the same?

When we standardize, we adjust for the size of the network:

```{r}

# Plot Network A using ggraph
p1s <- ggraph( graph_a, layout = layout_a ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_a )$color ), size = I( V( graph_a )$size ) ) ) +
  geom_node_text( aes( label = V( graph_a )$label_s ), size = 5 ) +
  labs( title = "Network A: Standardized Closeness Centrality" ) +
  theme_void()

# Plot Network b using ggraph
p2s <- ggraph( graph_b, layout = layout_b ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_b )$color ), size = I( V( graph_b )$size ) ) ) +
  geom_node_text( aes( label = V( graph_b )$label_s ), size = 5 ) +
  labs( title = "Network B: Standardized Closeness Centrality" ) +
  theme_void()

grid.arrange( p1s, nullGrob(), p2s, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

Now, apply the same thing to the rest of the network:

```{r}

# Plot Network A using ggraph
p1s <- ggraph( graph_a, layout = layout_a ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_a )$color ), size = I( V( graph_a )$size ) ) ) +
  geom_node_text( aes( label = V( graph_a )$label_s2 ), size = 5 ) +
  labs( title = "Network A: Standardized Closeness Centrality" ) +
  theme_void()

# Plot Network b using ggraph
p2s <- ggraph( graph_b, layout = layout_b ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_b )$color ), size = I( V( graph_b )$size ) ) ) +
  geom_node_text( aes( label = V( graph_b )$label_s2 ), size = 5 ) +
  labs( title = "Network B: Standardized Closeness Centrality" ) +
  theme_void()

grid.arrange( p1s, nullGrob(), p2s, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

### Closeness Centralization

In the prior chapter we discussed *graph centralization*, which measures the extent to which the nodes in a network differ from one another in their individual centrality scores. Put differently, how much variation is there in the distribution of centrality scores? As with degree centralization, we can can calculate the closeness centralization of the graph. A key difference, however, is that for closeness centralization, we use the standardized closeness score.

We can calculate closeness centralization as:

$$C_C = \frac{\sum\limits_{i=1}^g[C'_C(n^*)-C'_C(n_i)]}{[(g-2)(g-1)]/(2g-3)}$$

This may look imposing, but we can decompose it to see what it is doing:

-   $C'_C(n^*)$ is the largest standardized closeness score
-   $C'_C(n_i)$ is the standardized closeness score for node *i*
-   $\sum\limits_{i=1}^g[C'_C(n^*)-C'_C(n_i)]$ simply takes the sum of deviations for each node from the largest value
-   Finally, we take that sum and divide it by $[(g-2)(g-1)]/(2g-3)$

If I lost you in some of the math, don't worry. Let's calculate the closeness centralization score for our example graph:

```{r}

u_graph

```

Recall our table of standardized closeness scores:

| Node  | Distance | Closeness   | Standardized Closeness   |
| :---: | :------: | :---------: | :----------------------: |
| Jen   | 9        | 1/9 = 0.111 | 0.111*4 = 0.444          |
| Tom   | 6        | 1/6 = 0.167 | 0.167*4 = 0.668          |
| Bob   | 5        | 1/5 = 0.200 | 0.200*4 = 0.800          |
| Leaf  | 7        | 1/7 = 0.143 | 0.143*4 = 0.572          |
| Jim   | 7        | 1/7 = 0.143 | 0.143*4 = 0.572          |

*What is the largest standardized closeness score?* It is 0.8 for Bob. Also, we have $g = 5$. Now, we just plug these into our equation:

$$C_C = \frac{\sum\limits_{i=1}^g[0.800-C'_C(n_i)]}{[(5-2)(5-1)]/(2*5-3)}$$

| Node  | Distance | Closeness   | Standardized Closeness   | Deviations of Standardized Closeness $0.800-C'_C(n_i)$ |
| :---: | :------: | :---------: | :----------------------: | :----------------------------------------------------: |
| Jen   | 9        | 1/9 = 0.111 | 0.111*4 = 0.444          | 0.800 - 0.444 = 0.356 |
| Tom   | 6        | 1/6 = 0.167 | 0.167*4 = 0.668          | 0.800 - 0.668 = 0.132 |
| Bob   | 5        | 1/5 = 0.200 | 0.200*4 = 0.800          | 0.800 - 0.800 = 0.000 |
| Leaf  | 7        | 1/7 = 0.143 | 0.143*4 = 0.572          | 0.800 - 0.572 = 0.228 |
| Jim   | 7        | 1/7 = 0.143 | 0.143*4 = 0.572          | 0.800 - 0.572 = 0.228 |

Now, if we total all of the values in the last column we get 0.944. Plugging this in as our numerator, we get:

$$C_C = \frac{0.944}{[(5-2)(5-1)]/(2*5-3)} = \frac{0.944}{1.714} = 0.5507585 \approx 0.551$$

So, the closeness centralization score for our example graph is 0.551. *What does this mean?*

When closeness centrality is evenly dispersed, meaning that all nodes have the same closeness score, then the numerator in the equation will be zero and the quotient will be close to 0. When there is considerable inequality in the closeness centrality scores between nodes, the quotient will be closer to 1. Thus, closer to 1 indicates that the graph is hierarchically structured and closer to 0 means that the graph is more decentralized.

We can see this by examining two additional undirected networks:

```{r}
grid.arrange( star_graph, nullGrob(), circle_graph, ncol = 3, widths = c( 1, 0.2, 1 ) )
```

The closeness centralization score for the figure on the left is 1, whereas the closeness centralization score for the figure on the right is 0. 

## Directed Graphs

As was covered in the prior chapter, when we have a directed graph, we have to consider directionality when evaluating a centrality measure. For example, in the graph below, *how far is Jen from Tom?* *How far is Jen from Bob?*

```{r, adj-directed-define}

# create an adjacency matrix from the digraph
# note that you have to take the transpose for this to work correctly in the table
d_adj_matrix <- as_adjacency_matrix( t( digraph ), sparse = FALSE )

# melt the matrix into a long format suitable for ggplot
dfd <- melt( d_adj_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
dfd$Var2 <- factor( dfd$Var2, levels = rev( rownames( d_adj_matrix ) ) )

# plot the blank adjacency matrix
adj_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect( fill = "white", color = NA ),
        panel.grid.major = element_line( color = "gray" ),  
        panel.grid.minor = element_line( color = "lightgray" ),
        axis.text.x = element_text( hjust = 0.5, vjust = 0.5 ),
        axis.text.y = element_text( hjust = 0.5, vjust = 0.5 ),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Adjacency Matrix for Directed Graph" )

```

```{r}
d_graph
```

To answer this question, let's think again about what closeness measures. It is the inverse of the distance matrix. So, if we are looking at outgoing ties (i.e. $L(n_i,n_j)$) then our equation is the same for the undirected graph:

$$C_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} = \frac{1}{[\sum\limits_{j=1}^g d(n_i,n_j)]}$$

Note, however, that the distance of *i* to *j*, $d(n_i,n_j)$, does not necessarily equal the distance from *j* to *i*, $d(n_j,n_i)$. For example, the distance from Jen to Tom, $d(Jen,Tom)$, is 1. But, the distance from Tom to Jen, $d(Tom,Jen)$ is....? It is <u>undefined</u> because there is not an edge that originates from Tom and ends at Jen. This complicates things a bit. Let's work through an example to see what this looks like.

As before, we want to first build a *distance matrix* which shows the value of the geodesic (i.e. shortest path) from *i* to *j*.

```{r}

# create the distance matrix
d_matrix <- distances( digraph )

# melt the matrix into a long format suitable for ggplot
df_d <- melt( d_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
df_d$Var2 <- factor( dfd$Var2, levels = rev( rownames( d_adj_matrix ) ) )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

*How far is Jen from Tom?* *How far is Jen from Bob?* To figure this out, we can simply count the steps from Jen to Tom. Then Jen to Bob.

```{r}

# dummy the distances for Jen
df_d$Jen <- ifelse( rownames( df_d ) %in% c( 2:3 ), "show", "dont" )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

Jen is 1 step from Tom and 2 steps from Bob. *What about the rest of the scores for Jen?*

```{r}

# dummy the distances for Jen
df_d$Jen <- ifelse( rownames( df_d ) %in% c( 2:5 ), "show", "dont" )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

Now, recall that we sum these distances, and then take the reciprocal:

$$C_C(Jen) =  \frac{1}{[\sum\limits_{j=1}^g d(Jen,n_j)]} = \frac{1}{9} = 0.111$$

*What about Bob? How far is Bob from everyone?*


HERE: Stuck with trying to get the values correct here on the plot


```{r}

# dummy the distances for Bob
df_d$Bob <- ifelse( rownames( df_d ) %in% c( 11:12, 14:15 ), "show", "dont" )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Bob ), color = "black" ) +
  geom_text( aes( label = ifelse( Bob == "show", c( 1, 2, 3, 4 ), "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```




HERE WITH GOING THROUGH




## Test your Knowledge

-   ?

## Summary

In this chapter,..In the next chapter we will focus on a final conceptualization and operationalization of centrality: *betweenness*.
