```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE, 
  message=FALSE, 
  warning=FALSE, 
  fig.width = 10,
  fig.align = "center"
  )

library( here )      # for the directory
library( ggraph )    # for plotting
library( igraph )    # for working with graphs
library( ggplot2 )   # for plotting
library( gridExtra ) # for plotting multiple plots
library( grid )      # to include the null graphical object
library( reshape2 )  # for reshaping the matrix data for a plot

```

```{r, graphs-define}

# ----
# This creates the graphs for the plots below

# Create the graph objects
graph  <- graph_from_data_frame( 
  data.frame( 
    from = c( 1, 2, 3, 3, 4 ),
    to   = c( 2, 3, 4, 5, 5 ) ), 
  directed = FALSE )

digraph  <- graph_from_data_frame( 
  data.frame( 
    from = c( 1, 2, 3, 3, 4, 4, 5, 5  ),
    to   = c( 2, 3, 4, 5, 3, 5, 3, 4  ) ), 
  directed = TRUE )

# Set the random seed to render the same plot
set.seed( 507 )

# Set a fixed layout using the Fruchterman-Reingold layout
layout <- layout_with_fr( graph )
dilayout <- layout_with_fr( digraph )

# Set the labels
custom_labels <- c( "Jen","Tom","Bob","Leaf","Jim" )

# Assign the labels to the graph nodes
V( graph )$name <- custom_labels
V( digraph )$name <- custom_labels

```

```{r, plots-define}

# create the undirected graph
u_graph <-
  ggraph( graph, 
          layout = layout ) +                  
  geom_edge_link( color = "black", width = 0.8 ) +  
  geom_node_point( color = "#28a88d", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5 ) +
  ggtitle( "Undirected Graph" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
theme_void() 

# create the directed graph
d_graph <- 
  ggraph( digraph, 
          layout = dilayout ) +                  
  geom_node_point( color = "#fc23fc", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5) +
  geom_edge_link( aes( start_cap = label_rect( node1.name ), 
                     end_cap = circle( 5, 'mm' ) ),
                 arrow = arrow( length = unit( 0.02, "npc" ) ), 
                 color = "black", width = 0.8) +
  ggtitle( "Directed Graph" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  theme_void()  

```

# Network Description

Understanding the basic structure of a network is the first step toward making informed, data-driven decisions in crime analysis. Networks, whether simple or complex, contain key features that offer valuable insights into relationships, patterns, and points of vulnerability. As crime analysts, the ability to describe these features---such as the number of connections, the centrality of actors, and the overall shape of the network---is essential for identifying trends and prioritizing interventions. But how do we translate raw data into these insights?

By the end of the chapter, be able to answer this question:

-   How can I describe some simple features of a network?

## Simple Network Summary

So you have you network data, now what? The next step is to summarize its most critical features, which can guide your analysis. First, consider the size: How *big* is the network? This is the number of actors involved, giving a sense of its scale. Next, think about how tightly connected the network is—this is the *density*, or the proportion of actual connections relative to all possible ones. Another key question is separation: How fragmented or divided is the network? Understanding its *components* or *subgraphs* can highlight isolated clusters. Finally, consider the network's compactness, or its *diameter*—how far do actors have to travel through connections to reach others? These four basic measures lay the groundwork for describing and understanding the structure of any network you encounter in your analysis.

### Size

Put simply: how big is it? The size of the network is just the number of nodes in the network. We represent this with $g$. 

In the graphs below, what is $g$?

```{r}

# Display the plots side by side
grid.arrange( u_graph, nullGrob(), d_graph, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

### Density

Of the ties that could exist, what fraction are observed? This is the *density* of a graph: the number of edges compared to how many ties there could be.

#### Undirected Graph

The density of an **undirected** graph is given by: $$\frac{2L}{g(g-1)}$$ where $L$ is the number of edges and $g$ is the number of nodes. So, if a network has 5 nodes (i.e. $g$ is 5) then we could observe $g(g-1) = 5*(5-1) = 5*4 = 20$ edges. Take a look at the undirected graph below, what is the density?  

```{r}

u_graph

```

First we identify the quantities: $L$ is 5 and $g$ is 5. Now, let's just plug through the math: $$\frac{2L}{g(g-1)} = \frac{2*5}{5*(5-1)} = \frac{10}{20} = 0.5$$ So, 0.5, or half, of the ties that could occur in this network did occur.

#### Directed Graph

The density of a **directed** graph is given by: $$\frac{L}{g(g-1)}$$ where $L$ is the number of edges and $g$ is the number of nodes. So, if a network has 5 nodes (i.e. $g$ is 5) then we could observe $g(g-1) = 5*(5-1) = 5*4 = 20$ edges. Take a look at the directed graph below, what is the density?  

```{r}

d_graph

```

First we identify the quantities: $L$ is 8 and $g$ is 5. Why is $L$ 8? Since the graph is directed, a dyad (or two nodes) can have two ties. For example, there is a tie from Jim to Leaf and a tie from Leaf to Jim. If we count up all the arrowheads, we get to 8.Now, let's just plug through the math: $$\frac{L}{g(g-1)} = \frac{8}{5*(5-1)} = \frac{8}{20} = 0.4$$ So, 0.4 of the ties that could occur in this network did occur.

HERE WITH COMPONENTS

## Test your Knowledge

-   What was the key question behind the Konigsberg bridge problem? How did Euler represent this problem using points and lines? What was the significance of this representation?

-   How is a graph represented in terms of nodes and edges? What does the notation $G = (N,L)$ mean?

-   What does it mean for two nodes to be **adjacent** in a graph?

-   In an undirected graph, how are the connections between nodes represented? How is this different from a directed graph?

-   In an adjacency matrix, what does the value in the $i^{th}$ row and $j^{th}$ column represent?

-   What is the difference between an **undirected** and a **directed** adjacency matrix? In an **undirected** graph, is the $i^{th}$ row equal to the $j^{th}$ column? What about for a **directed** graph?

-   Define the following terms as they relate to matrices: Scalar, Column vector, Row vector, Matrix, Square matrix, Order of a matrix.

## Summary

This chapter has how to represent networks using matrices. In the next chapter, we will start looking at how to analyze these matrices.
