```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE, 
  message=FALSE, 
  warning=FALSE, 
  fig.width = 10,
  fig.align = "center"
  )

library( here )      # for the directory
library( ggraph )    # for plotting
library( igraph )    # for working with graphs
library( ggplot2 )   # for plotting
library( gridExtra ) # for plotting multiple plots
library( grid )      # to include the null graphical object
library( reshape2 )  # for reshaping the matrix data for a plot

```

# Betweenness Centrality

In the prior chapter, we examined the idea of "centrality" for nodes and "centralization" for graphs, focusing attention on the conceptual aspects of what it means to be *central* as well as the ways in which we can operationalize the concept using *degree centrality* and *closeness centrality*. In this chapter we want to continue thinking about what it means to be central, but through a different conceptual lens: **betweenness**. Betweenness centrality serves as a powerful indicator of actors who bridge otherwise disconnected parts of a network, offering unique control over the flow of information and resources across network boundaries. Unlike *degree* centrality, which emphasizes direct ties, or *closeness* centrality, which identifies actors closest to others, betweenness centrality highlights those who can influence inter-group dynamics. A compelling example is seen in a study by @calderoni2014social, which examined the Italian Mafia network, specifically the Ndrangheta organization. Calderoni found that individuals with high betweenness centrality played pivotal roles in connecting local criminal cells to broader, often international networks. These "connectors" were essential for coordinating illicit activities across regions and facilitating transactions that would otherwise remain confined to isolated groups. For law enforcement, targeting these high-betweenness actors provided a strategic leverage point to disrupt cross-regional collaboration, making them critical targets in efforts to dismantle organized crime networks. By understanding the role of betweenness, crime analysts can better assess how key individuals maintain structural cohesion within social networks.

By the end of this chapter, you should be able to:

-   Explain the conceptualization of "centrality" as it pertains to *betweenness centrality*.
-   Calculate betweenness centrality scores for a set of nodes.
-   Calculate a betweenness centralization score for a graph.

```{r, graphs-define}

# ----
# This creates the graphs for the plots below

# Create the graph objects
ugraph  <- graph_from_data_frame( 
  data.frame( 
    from = c( 1, 2, 3, 3, 4 ),
    to   = c( 2, 3, 4, 5, 5 ) ), 
  directed = FALSE )

digraph  <- graph_from_data_frame( 
  data.frame( 
    from = c( 1, 2, 3, 3, 4, 4, 5, 5  ),
    to   = c( 2, 3, 4, 5, 3, 5, 3, 4  ) ), 
  directed = TRUE )

star_net <- graph_from_data_frame(
  data.frame(
    from = c( 1, 1, 1, 1 ),
    to   = c( 2, 3, 4, 5 ) ),
  directed = FALSE )

circle_net <- graph_from_data_frame(
  data.frame(
    from = c( 1, 2, 3, 4, 5 ),
    to   = c( 2, 3, 4, 5, 1 ) ),
  directed = FALSE )

# Set the random seed to render the same plot
set.seed( 507 )

# Set a fixed layout using the Fruchterman-Reingold layout
layout <- layout_with_fr( ugraph )
dilayout <- layout_with_fr( digraph )

starlayout <- layout_with_fr( star_net )
circlelayout <- layout_with_fr( circle_net )


# Set the labels
custom_labels <- c( "Jen","Tom","Bob","Leaf","Jim" )

# Assign the labels to the graph nodes
V( ugraph )$name <- custom_labels
V( digraph )$name <- custom_labels

```

```{r, plots-define}

# create the undirected plots
u_graph <-
  ggraph( ugraph, 
          layout = layout ) +                  
  geom_edge_link( color = "black", width = 0.8 ) +  
  geom_node_point( color = "#28a88d", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5 ) +
  ggtitle( "Undirected Graph" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
theme_void() 

star_graph <-
  ggraph( star_net, 
          layout = starlayout ) +                  
  geom_edge_link( ) +  
  geom_node_point( color = "#28a88d", size = 10 ) +
  ggtitle( "Graph w/ High Degree Centralization" ) + 
theme_void() 

circle_graph <-
  ggraph( circle_net, 
          layout = circlelayout ) +                  
  geom_edge_link( ) +  
  geom_node_point( color = "#28a88d", size = 10 ) +
  ggtitle( "Graph w/ Low Degree Centralization" ) + 
theme_void() 


# create the directed graphs
d_graph <- 
  ggraph( digraph, 
          layout = dilayout ) +                  
  geom_node_point( color = "#fc23fc", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5) +
  geom_edge_link( aes( start_cap = label_rect( node1.name ), 
                     end_cap = circle( 5, 'mm' ) ),
                 arrow = arrow( length = unit( 0.02, "npc" ) ), 
                 color = "black", width = 0.8) +
  ggtitle( "Directed Graph" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  theme_void()  

```

## Case Study: Money-Laundering and Betweenness

In a case study examining the role of betweenness centrality in criminal networks, @malm2011using analyzed money-laundering to understand how different actors influenced network cohesion and resource flow. Analyzing data from intelligence reports, the authors found that the majority of money-laundering in this drug market was self-laundering (where individuals laundered their own illicit earnings), rather than employing specialized professionals. Using betweenness centrality to assess how well-positioned launderers were within the network, @malm2011using found that self-launderers generally exhibited higher betweenness centrality, showing they often acted as key brokers facilitating information flow and transactions across different parts of the network. The findings suggested that targeting individuals with high betweenness centrality, particularly those involved in self-laundering, may be most effective for disrupting drug networks.

## Betweenness Centrality Definition

*Betweenness* centrality is based on the number of <u>shortest</u> paths between two nodes, *j* and *k*, that node *i* resides on. Recall that a **path** is an edge or sequence of edges that connect two nodes and a **geodesic** is the shortest path between two nodes.

## Undirected Graphs

```{r, adj-undirected-define}

# create an adjacency matrix from the graph
adj_matrix <- as_adjacency_matrix( ugraph, sparse = FALSE )

# melt the matrix into a long format suitable for ggplot
df <- melt( adj_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
df$Var2 <- factor( df$Var2, levels = rev( rownames( adj_matrix ) ) )

# plot the adjacency matrix
adj_to_plot <- ggplot( df, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  geom_text( aes( label = value ), color = "black" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Adjacency Matrix for Undirected Graph" )

```

### Betweenness Centrality

For an undirected binary graph, betweenness centrality is:

$$C_B(n_i) = \sum\limits_{j<k} g_{jk}(n_i) / g_{jk}$$

This equation is a bit different from those for degree and closeness centrality. Let's unpack it:

-   The denominator, $g_{jk}$, is the number of geodesics linking *j* to *k*.
-   The numerator, $g_{jk}(n_i)$, is the number of geodesics linking *j* and *k* that contain *i*.\
-   Betweenness centrality is the ratio of the geodesics between *j* and *k* that contain *i*.

In words, if *j* has to go through *i* to reach *k*, $j-i-k$, then *i* will have high betweenness because *i* is **between** *j* and *k*.

Let's consider an example. Take a look at this graph:

```{r}
u_graph
```

*What are the* <u>paths</u> between Jen and Jim?

There are two:

-   Jen to Tom to Bob to Jim
-   Jen to Tom to Bob to Leaf to Jim

We can see this visually:

```{r}

# set up the colors
edge_colors1 <- c("#28a88d", "#28a88d", "grey80", "#28a88d", "grey80" )
edge_colors2 <- c("#28a88d", "#28a88d", "#28a88d", "grey80", "#28a88d" )

E( ugraph )$color1 <- edge_colors1
E( ugraph )$color2 <- edge_colors2


# do the plots
plot1 <-
  ggraph( ugraph, 
          layout = layout ) +                  
  geom_edge_link( aes( color = edge_colors1 ), width = 0.8, show.legend = FALSE ) +  
  geom_node_point( color = "#28a88d", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5 ) +
  ggtitle( "Jen to Tom to Bob to Jim" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_edge_color_identity() +
theme_void() 

plot2 <-
  ggraph( ugraph, 
          layout = layout ) +                  
  geom_edge_link( aes( color = edge_colors2 ), width = 0.8, show.legend = FALSE ) +  
  geom_node_point( color = "#28a88d", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5 ) +
  ggtitle( "Jen to Tom to Bob to Leaf to Jim" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_edge_color_identity() +
theme_void() 

# now plot the plots
grid.arrange( plot1, plot2, ncol = 2 )

```

Note that Bob is on both of those paths. He is *between* Jen and Jim on both paths. Note, however, that only one of the paths is a <u>geodesic path</u> (the Jen-\>Tom-\>Bob-\>Jim path). That is because it is shorter.

Now, let's calculate the betweenness centrality for Bob. The **first** thing we need to find are the geodesic paths. Then, figure out how many of those geodesic paths contain Bob. We do this be creating a table of geodesic proportions for Bob.

#### Geodesic Proportions for *Bob*

|      | Jen | Tom | Leaf |              Jim              |
|:-----|:---:|:---:|:----:|:-----------------------------:|
| Jen  |     |     |      | [/?]{style="color: #28a88d;"} |
| Tom  |     |     |      |                               |
| Leaf |     |     |      |                               |
| Jim  |     |     |      |                               |

*How many geodesics from Jen to Jim?* There is 1 geodesic from **Jen** to **Jim**. So, we would just add that to our table:

#### Geodesic Proportions for *Bob*

|      | Jen | Tom | Leaf |              Jim              |
|:-----|:---:|:---:|:----:|:-----------------------------:|
| Jen  |     |     |      | [/1]{style="color: #28a88d;"} |
| Tom  |     |     |      |                               |
| Leaf |     |     |      |                               |
| Jim  |     |     |      |                               |

Now, how many geodesics from **Jen** to **Jim** include **Bob**? Bob is <u>on</u> the *only* geodesic from Jen to Jim. So, we would just add that to our table:

#### Geodesic Proportions for *Bob*

|      | Jen | Tom | Leaf |              Jim               |
|:-----|:---:|:---:|:----:|:------------------------------:|
| Jen  |     |     |      | [1/1]{style="color: #28a88d;"} |
| Tom  |     |     |      |                                |
| Leaf |     |     |      |                                |
| Jim  |     |     |      |                                |

What about Jen to Tom?

```{r}

edge_colors1 <- c("#28a88d", "grey80", "grey80", "grey80", "grey80" )
E( ugraph )$color1 <- edge_colors1

  ggraph( ugraph, 
          layout = layout ) +                  
  geom_edge_link( aes( color = edge_colors1 ), width = 0.8, show.legend = FALSE ) +  
  geom_node_point( color = "#28a88d", size = 15 ) +
  geom_node_text( aes( label = name ), 
                 color = "black", 
                 size = 5,        
                 vjust = 0.5,     
                 hjust = 0.5 ) +
  ggtitle( "Jen to Tom" ) + 
  scale_x_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_y_continuous( expand = expansion( mult = c( 0.2, 0.2 ) ) ) +
  scale_edge_color_identity() +
theme_void() 

```

#### Geodesic Proportions for *Bob*

|      | Jen |              Tom               | Leaf | Jim |
|:-----|:---:|:------------------------------:|:----:|:---:|
| Jen  |     | [?/?]{style="color: #28a88d;"} |      | 1/1 |
| Tom  |     |                                |      |     |
| Leaf |     |                                |      |     |
| Jim  |     |                                |      |     |

There is 1 geodesic from Jen to Tom, <u>and</u> Bob is not on it. So, that is a 0 for the numerator. We can update the table accordingly:

#### Geodesic Proportions for *Bob*

|      | Jen |              Tom               | Leaf | Jim |
|:-----|:---:|:------------------------------:|:----:|:---:|
| Jen  |     | [0/1]{style="color: #28a88d;"} |      | 1/1 |
| Tom  |     |                                |      |     |
| Leaf |     |                                |      |     |
| Jim  |     |                                |      |     |

Finally, what about Jen to Leaf?

#### Geodesic Proportions for *Bob*

|      | Jen | Tom |              Leaf              | Jim |
|:-----|:---:|:---:|:------------------------------:|:---:|
| Jen  |     | 0/1 | [?/?]{style="color: #28a88d;"} | 1/1 |
| Tom  |     |     |                                |     |
| Leaf |     |     |                                |     |
| Jim  |     |     |                                |     |

Bob is on the geodesic from Jen to Leaf:

|      | Jen | Tom |              Leaf              | Jim |
|:-----|:---:|:---:|:------------------------------:|:---:|
| Jen  |     | 0/1 | [1/1]{style="color: #28a88d;"} | 1/1 |
| Tom  |     |     |                                |     |
| Leaf |     |     |                                |     |
| Jim  |     |     |                                |     |

Of the geodesics (Jen, Tom; Jen, Leaf; and Jen, Jim) how many include Bob? It is 2.

|      | Jen | Tom | Leaf | Jim |
|:-----|:---:|:---:|:----:|:---:|
| Jen  |     | 0/1 | 1/1  | 1/1 |
| Tom  |     |     |      |     |
| Leaf |     |     |      |     |
| Jim  |     |     |      |     |


HERE



```{r}

# create the distance matrix
d_matrix <- distances( ugraph )

# melt the matrix into a long format suitable for ggplot
dfd <- melt( d_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
dfd$Var2 <- factor( df$Var2, levels = rev( rownames( adj_matrix ) ) )

# plot the blank distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

For our graph, we want to first build a *distance matrix* which shows the value of the geodesic (i.e. shortest path) from *i* to *j*. We can do this by looking at each node and seeing the fewest steps that are needed to reach another node. We can see that Jen can reach Tom in 1 step (i.e. $L_{Jen,Tom}$) and that Jen can reach Bob in 2 steps (i.e. $L_{Jen,Tom}$ & $L_{Tom,Bob}$).

```{r}

# dummy the distances for Jen
dfd$Jen <- ifelse( rownames( dfd ) %in% c( 1:5 ), "show", "dont" )

# plot the distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#28a88d", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

Since the graph is undirected, the distance matrix is symmetric about the diagonal (just like we saw with an adjacency matrix). That is, $d(n_i,n_j) = d(n_j,n_i)$. This gives us:

```{r}

# dummy the distances for Jen
dfd$Jen <- ifelse( rownames( dfd ) %in% c( 1:5, 6, 11, 16, 21 ), "show", "dont" )

# plot the distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#28a88d", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

Now, what about the rest of the distance matrix? If we follow what we did above for everyone else, then we get the following distance matrix:

```{r}

# plot the distance matrix
d_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  geom_text( aes( label = value ), color = "black" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Undirected Graph" )

# plot the network and the distance matrix
grid.arrange( u_graph, d_to_plot, ncol = 2 )

```

Now, it we take the sum of the rows, this gives us $\sum\limits_{j=1}^g d(n_i,n_j)$. Note that since the graph is undirected, the column sum would give us the same values. Since we have $\sum\limits_{j=1}^g d(n_i,n_j)$, all we need to do now is take the reciprocal by dividing 1 by $\sum\limits_{j=1}^g d(n_i,n_j)$:

$$C_C(n_i) =  \frac{1}{[\sum\limits_{j=1}^g d(n_i,n_j)]}$$

That gives us the closeness centrality score for each individual:

| Node | Distance |  Closeness  |
|:----:|:--------:|:-----------:|
| Jen  |    9     | 1/9 = 0.111 |
| Tom  |    6     | 1/6 = 0.167 |
| Bob  |    5     | 1/5 = 0.200 |
| Leaf |    7     | 1/7 = 0.143 |
| Jim  |    7     | 1/7 = 0.143 |

### Standardization

As discussed in the prior chapter, centrality measures are sensitive to the size of the graph, *g*. In the case of *closeness* centrality, having to sum over more nodes will push scores closer to zero. Think about it: if we added more rows and columns to our distance matrix, that will increase the denominator, making the quotient smaller. This means that we can't compare closeness centrality scores across graphs of different sizes. *Solution*?

Standardize! If we want to compare nodes in different sized networks, then we just take into account the number of nodes and the maximum possible nodes to which *i* could be connected. In an undirected graph, this is *g-1*.

Thus, we can calculate a standardized closeness centrality for an undirected graph as:

$$C'_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} \times [g-1]$$

For this graph, the denominator, *g-1* is $5 - 1 = 4$. Plugging that in we get: $C'_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} \times 4$.

| Node | Distance |  Closeness  | Standardized Closeness |
|:----:|:--------:|:-----------:|:----------------------:|
| Jen  |    9     | 1/9 = 0.111 |    0.111\*4 = 0.444    |
| Tom  |    6     | 1/6 = 0.167 |    0.167\*4 = 0.668    |
| Bob  |    5     | 1/5 = 0.200 |    0.200\*4 = 0.800    |
| Leaf |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |
| Jim  |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |

Standardization is particularly important with *closeness* centrality because of the sensitivity of the measure to the size of the network. Take for example, the following networks:

```{r}

# set up the objects for the plot
graph_a <- graph_from_adjacency_matrix( 
  matrix(
    c( 0, 1, 0,
      1, 0, 1,
      0, 1, 0),
    byrow = TRUE, ncol = 3 ), 
  mode = "undirected" 
  )

graph_b <- graph_from_adjacency_matrix( 
  matrix(
    c( 0, 1, 1, 1,
      1, 0, 1, 1,
      1, 1, 0, 1,
      1, 1, 1, 0 ),
    byrow = TRUE, ncol = 4 ), 
  mode = "undirected" 
)

# set the vertex and edge properties for Network A
V( graph_a )$color <- c( "lightblue", "grey80", "grey80" )
V( graph_a )$size  <- 10
V( graph_a )$label <- c( "0.33", "", "" )
V( graph_a )$label_s <- c( "0.66","","" )
V( graph_a )$label_s2 <- c( "0.66","1","0.66" )
E( graph_a )$color <- "grey50"

# set the vertex and edge properties for Network B
V( graph_b )$color <- c( "orange", "grey80", "grey80", "grey80" )
V( graph_b )$size <- 10
V( graph_b )$label <- c( "0.33", "", "", "" )
V( graph_b )$label_s <- c( "1", "", "", "" )
V( graph_b )$label_s2 <- c( "1", "1", "1", "1" )
E( graph_b )$color <- "grey50"

# set the seed for the layout
set.seed( 507 ) 

# set the layouts
layout_a <- layout_with_fr( graph_a )
layout_b <- layout_with_fr( graph_b )

```

```{r}

# Plot Network A using ggraph
p1 <- ggraph( graph_a, layout = layout_a ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_a )$color ), size = I( V( graph_a )$size ) ) ) +
  geom_node_text( aes( label = V( graph_a )$label ), size = 5 ) +
  labs( title = "Network A: Raw Closeness Centrality" ) +
  theme_void()

# Plot Network b using ggraph
p2 <- ggraph( graph_b, layout = layout_b ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_b )$color ), size = I( V( graph_b )$size ) ) ) +
  geom_node_text( aes( label = V( graph_b )$label ), size = 5 ) +
  labs( title = "Network B: Raw Closeness Centrality" ) +
  theme_void()

grid.arrange( p1, nullGrob(), p2, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

The plot above shows two different graphs with the *raw* closeness centrality scores for two nodes. Note that they have the exact same score. Yet, would you say they are equally close to other nodes in the network? In other words, that their closeness to other nodes is the same?

When we standardize, we adjust for the size of the network:

```{r}

# Plot Network A using ggraph
p1s <- ggraph( graph_a, layout = layout_a ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_a )$color ), size = I( V( graph_a )$size ) ) ) +
  geom_node_text( aes( label = V( graph_a )$label_s ), size = 5 ) +
  labs( title = "Network A: Standardized Closeness Centrality" ) +
  theme_void()

# Plot Network b using ggraph
p2s <- ggraph( graph_b, layout = layout_b ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_b )$color ), size = I( V( graph_b )$size ) ) ) +
  geom_node_text( aes( label = V( graph_b )$label_s ), size = 5 ) +
  labs( title = "Network B: Standardized Closeness Centrality" ) +
  theme_void()

grid.arrange( p1s, nullGrob(), p2s, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

Now, apply the same thing to the rest of the network:

```{r}

# Plot Network A using ggraph
p1s <- ggraph( graph_a, layout = layout_a ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_a )$color ), size = I( V( graph_a )$size ) ) ) +
  geom_node_text( aes( label = V( graph_a )$label_s2 ), size = 5 ) +
  labs( title = "Network A: Standardized Closeness Centrality" ) +
  theme_void()

# Plot Network b using ggraph
p2s <- ggraph( graph_b, layout = layout_b ) +
  geom_edge_link( aes( edge_alpha = 0.5 ), color = "grey50", show.legend = FALSE ) +
  geom_node_point( aes( color = I( V( graph_b )$color ), size = I( V( graph_b )$size ) ) ) +
  geom_node_text( aes( label = V( graph_b )$label_s2 ), size = 5 ) +
  labs( title = "Network B: Standardized Closeness Centrality" ) +
  theme_void()

grid.arrange( p1s, nullGrob(), p2s, ncol = 3, widths = c( 1, 0.2, 1 ) )

```

### Closeness Centralization

In the prior chapter we discussed *graph centralization*, which measures the extent to which the nodes in a network differ from one another in their individual centrality scores. Put differently, how much variation is there in the distribution of centrality scores? As with degree centralization, we can can calculate the closeness centralization of the graph. A key difference, however, is that for closeness centralization, we use the standardized closeness score.

We can calculate closeness centralization as:

$$C_C = \frac{\sum\limits_{i=1}^g[C'_C(n^*)-C'_C(n_i)]}{[(g-2)(g-1)]/(2g-3)}$$

This may look imposing, but we can decompose it to see what it is doing:

-   $C'_C(n^*)$ is the largest standardized closeness score
-   $C'_C(n_i)$ is the standardized closeness score for node *i*
-   $\sum\limits_{i=1}^g[C'_C(n^*)-C'_C(n_i)]$ simply takes the sum of deviations for each node from the largest value
-   Finally, we take that sum and divide it by $[(g-2)(g-1)]/(2g-3)$

If I lost you in some of the math, don't worry. Let's calculate the closeness centralization score for our example graph:

```{r}

u_graph

```

Recall our table of standardized closeness scores:

| Node | Distance |  Closeness  | Standardized Closeness |
|:----:|:--------:|:-----------:|:----------------------:|
| Jen  |    9     | 1/9 = 0.111 |    0.111\*4 = 0.444    |
| Tom  |    6     | 1/6 = 0.167 |    0.167\*4 = 0.668    |
| Bob  |    5     | 1/5 = 0.200 |    0.200\*4 = 0.800    |
| Leaf |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |
| Jim  |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |

*What is the largest standardized closeness score?* It is 0.8 for Bob. Also, we have $g = 5$. Now, we just plug these into our equation:

$$C_C = \frac{\sum\limits_{i=1}^g[0.800-C'_C(n_i)]}{[(5-2)(5-1)]/(2*5-3)}$$

| Node | Distance |  Closeness  | Standardized Closeness | Deviations of Standardized Closeness $0.800-C'_C(n_i)$ |
|:-------------:|:-------------:|:-------------:|:-------------:|:-------------:|
| Jen  |    9     | 1/9 = 0.111 |    0.111\*4 = 0.444    |                 0.800 - 0.444 = 0.356                  |
| Tom  |    6     | 1/6 = 0.167 |    0.167\*4 = 0.668    |                 0.800 - 0.668 = 0.132                  |
| Bob  |    5     | 1/5 = 0.200 |    0.200\*4 = 0.800    |                 0.800 - 0.800 = 0.000                  |
| Leaf |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |                 0.800 - 0.572 = 0.228                  |
| Jim  |    7     | 1/7 = 0.143 |    0.143\*4 = 0.572    |                 0.800 - 0.572 = 0.228                  |

Now, if we total all of the values in the last column we get 0.944. Plugging this in as our numerator, we get:

$$C_C = \frac{0.944}{[(5-2)(5-1)]/(2*5-3)} = \frac{0.944}{1.714} = 0.5507585 \approx 0.551$$

So, the closeness centralization score for our example graph is 0.551. *What does this mean?*

When closeness centrality is evenly dispersed, meaning that all nodes have the same closeness score, then the numerator in the equation will be zero and the quotient will be close to 0. When there is considerable inequality in the closeness centrality scores between nodes, the quotient will be closer to 1. Thus, closer to 1 indicates that the graph is hierarchically structured and closer to 0 means that the graph is more decentralized.

We can see this by examining two additional undirected networks:

```{r}
grid.arrange( star_graph, nullGrob(), circle_graph, ncol = 3, widths = c( 1, 0.2, 1 ) )
```

The closeness centralization score for the figure on the left is 1, whereas the closeness centralization score for the figure on the right is 0.

## Directed Graphs

As was covered in the prior chapter, when we have a directed graph, we have to consider directionality when evaluating a centrality measure. For example, in the graph below, *how far is Jen from Tom?* *How far is Jen from Bob?*

```{r, adj-directed-define}

# create an adjacency matrix from the digraph
# note that you have to take the transpose for this to work correctly in the table
d_adj_matrix <- as_adjacency_matrix( t( digraph ), sparse = FALSE )

# melt the matrix into a long format suitable for ggplot
dfd <- melt( d_adj_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
dfd$Var2 <- factor( dfd$Var2, levels = rev( rownames( d_adj_matrix ) ) )

# plot the blank adjacency matrix
adj_to_plot <- ggplot( dfd, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect( fill = "white", color = NA ),
        panel.grid.major = element_line( color = "gray" ),  
        panel.grid.minor = element_line( color = "lightgray" ),
        axis.text.x = element_text( hjust = 0.5, vjust = 0.5 ),
        axis.text.y = element_text( hjust = 0.5, vjust = 0.5 ),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Adjacency Matrix for Directed Graph" )

```

```{r}
d_graph
```

To answer this question, let's think again about what closeness measures. It is the inverse of the distance matrix. So, if we are looking at outgoing ties (i.e. $L(n_i,n_j)$) then our equation is the same for the undirected graph:

$$C_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} = \frac{1}{[\sum\limits_{j=1}^g d(n_i,n_j)]}$$

Note, however, that the distance of *i* to *j*, $d(n_i,n_j)$, does not necessarily equal the distance from *j* to *i*, $d(n_j,n_i)$. For example, the distance from Jen to Tom, $d(Jen,Tom)$, is 1. But, the distance from Tom to Jen, $d(Tom,Jen)$ is....? It is <u>undefined</u> because there is not an edge that originates from Tom and ends at Jen. This complicates things a bit. Let's work through an example to see what this looks like.

### Outdegree

Closeness centrality in directed graphs usually focuses on the send network (i.e. outgoing ties). Remember, *row* sums are the *outdegree*, so row distance tells you how close <u>you</u> are to others. That is, how many steps it takes <u>you</u> to reach another node.

As before, we want to first build a *distance matrix* which shows the value of the geodesic (i.e. shortest path) from *i* to *j*.

```{r}

# create the distance matrix
d_matrix <- distances( digraph )

# melt the matrix into a long format suitable for ggplot
df_d <- melt( d_matrix )

# Reverse the order of the y-axis labels so they aren't backwards
df_d$Var2 <- factor( dfd$Var2, levels = rev( rownames( d_adj_matrix ) ) )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

*How far is Jen from Tom?* *How far is Jen from Bob?* To figure this out, we can simply count the steps from Jen to Tom. Then Jen to Bob.

```{r}

# dummy the distances for Jen
df_d$Jen <- ifelse( rownames( df_d ) %in% c( 2:3 ), "show", "dont" )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

Jen is 1 step from Tom and 2 steps from Bob. *What about the rest of the scores for Jen?*

```{r}

# dummy the distances for Jen
df_d$Jen <- ifelse( rownames( df_d ) %in% c( 2:5 ), "show", "dont" )

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Jen ), color = "black" ) +
  geom_text( aes( label = ifelse( Jen == "show", value, "" ) ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

Now, recall that we sum these distances, and then take the reciprocal:

$$C_C(Jen) =  \frac{1}{[\sum\limits_{j=1}^g d(Jen,n_j)]} = \frac{1}{9} = 0.111$$

*What about Bob? How far is Bob from everyone?*

```{r}

# dummy the distances for Bob
df_d$Bob <- ifelse( rownames( df_d ) %in% c( 11:12, 14:15 ), "show", "dont" )

# create a vector for the labels
custom_labels <- c("?", "?", "?", "?")

# assign these to show in the plot
df_d$label <- ifelse( df_d$Bob == "show", 
                     custom_labels[ match( rownames( df_d ), c( 11, 12, 14, 15 ) ) ], 
                     "")

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Bob ), color = "black" ) +
  geom_text( aes( label = label ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

How far is Bob from Jen? Bob, is "infinite distance" from Jen because there is not a directed path by when he can reach her. This is also the case for the distance from Bob to Tom. But, Bob does have directed ties to Jim and Leaf. With directionality in mind, they are unreachable. But, Bob is 1 step from Leaf and 1 step from Jim along a directed path:

```{r}

# dummy the distances for Bob
df_d$Bob <- ifelse( rownames( df_d ) %in% c( 11:12, 14:15 ), "show", "dont" )

# create a vector for the labels
custom_labels <- c(Inf, Inf, 1, 1)

# assign these to show in the plot
df_d$label <- ifelse( df_d$Bob == "show", 
                     custom_labels[ match( rownames( df_d ), c( 11, 12, 14, 15 ) ) ], 
                     "")

# plot the blank distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
    geom_tile( aes( fill = Bob ), color = "black" ) +
  geom_text( aes( label = label ), color = "black" ) +
  scale_fill_manual(values = c("show" = "#fc23fc", "dont" = "white"), guide = "none") +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

So what is Bob's closeness centrality score? If we do not count the cells that have "infinite distance", and only count cells where there are values, then we just count the distance for Leaf and Jim, which sums to 2. We just plug this in:

$$C_C(Bob) =  \frac{1}{[\sum\limits_{j=1}^g d(Bob,n_j)]} = \frac{1}{2} = 0.500$$

Note that this changes the denominator for each node.

If we fill in our matrix, then we have the following distance values:

```{r}

# plot the distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  geom_text( aes( label = as.vector( t( distances( digraph, mode = "out" ) ) ) ) ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph (Outdegree)" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

Going through the rest of the matrix, we get a closeness centrality score for each individual:

| Node | Distance |  Closeness  |
|:----:|:--------:|:-----------:|
| Jen  |    9     | 1/9 = 0.111 |
| Tom  |    5     | 1/5 = 0.200 |
| Bob  |    2     | 1/2 = 0.500 |
| Leaf |    2     | 1/2 = 0.500 |
| Jim  |    2     | 1/2 = 0.500 |

#### Standardization

Don't, forget: raw centrality scores are sensitive to the size of the graph. So, as before, we can standardize our closeness score as:

$$C'_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} \times [g-1]$$

| Node | Distance |  Closeness  | Standardized Closeness |
|:----:|:--------:|:-----------:|:----------------------:|
| Jen  |    9     | 1/9 = 0.111 |    0.111\*4 = 0.444    |
| Tom  |    5     | 1/5 = 0.200 |    0.200\*4 = 0.800    |
| Bob  |    2     | 1/2 = 0.500 |    0.500\*4 = 0.200    |
| Leaf |    2     | 1/2 = 0.500 |    0.500\*4 = 0.200    |
| Jim  |    2     | 1/2 = 0.500 |    0.500\*4 = 0.200    |

### Indegree

Note that in the prior section we were looking at *outdegree*. When we are looking at *outgoing* ties the closeness scores reflect how close a <u>you</u> are to other nodes. For example, Tom's distance to Leaf is 2, meaning that it takes 2 outgoing ties to reach Leaf. What about the other way around? We can reverse this and look at the *incoming* ties by focusing on *indegree*. Remember, *column* sums are the *indegree*, so column distance tells you how close <u>others</u> are to you. That is, how many steps it takes a <u>node</u> to reach you. The difference is a bit nuanced, but meaningful. In this case, instead of summing the rows as we did above, we could sum the columns.

```{r}

# plot the distance matrix
d_to_plot <- ggplot( df_d, aes( Var1, Var2 ) ) +
  geom_tile( color = "black", fill = "white" ) +
  geom_text( aes( label = as.vector( distances( digraph, mode = "out" ) ) ) ) +
  coord_fixed() +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "white", color = NA),
        panel.grid.major = element_line(color = "gray"),  
        panel.grid.minor = element_line(color = "lightgray"),
        axis.text.x = element_text(hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(hjust = 0.5, vjust = 0.5),
        axis.title = element_blank(),
        axis.text.x.top = element_text(hjust = 0.5, vjust = 0.5),
        axis.ticks = element_blank()) + 
  labs(x = NULL, y = NULL) +  
  scale_y_discrete(position = "left") +
  scale_x_discrete(position = "top")  +
  ggtitle( "Distance Matrix for Directed Graph (Indegree)" )

# plot the network and the distance matrix
grid.arrange( d_graph, d_to_plot, ncol = 2 )

```

If we sum the columns, then we get these scores:

| Node | Distance |    Closeness    |
|:----:|:--------:|:---------------:|
| Jen  |    0     | 1/0 = Undefined |
| Tom  |    1     |   1/1 = 1.000   |
| Bob  |    5     |   1/5 = 0.200   |
| Leaf |    7     |   1/7 = 0.143   |
| Jim  |    7     |   1/7 = 0.143   |

#### Standardization

We can then standardize our closeness score as:

$$C'_C(n_i) = [\sum\limits_{j=1}^g d(n_i,n_j)]^{-1} \times [g-1]$$

| Node | Distance |    Closeness    |  Standardized Closeness  |
|:----:|:--------:|:---------------:|:------------------------:|
| Jen  |    0     | 1/0 = Undefined | Undefined\*4 = Undefined |
| Tom  |    1     |   1/1 = 1.000   |     1.000\*4 = 4.000     |
| Bob  |    5     |   1/5 = 0.200   |     0.200\*4 = 0.800     |
| Leaf |    7     |   1/7 = 0.143   |     0.143\*4 = 0.571     |
| Jim  |    7     |   1/7 = 0.143   |     0.143\*4 = 0.571     |

### Conversion to Undirected Graph

**NOTE**: While these are correct, mathematically, they are far less intuitive than the scores we calculated based on the *outgoing* ties. For example, Tom's closeness score is 1. But, Bob, Leaf, and Jim cannot reach him. Also, note that Tom's standardized score is greater than 1. For these reasons, we often convert the directed graph to an undirected graph and calculate the scores on the undirected graph. In this case, the directed graph we have here would revert to the undirected graph we examined above in the section on undirected graphs.

### Closeness Centralization

The denominator for group closeness centralization is not defined for directed graphs. To get a centralization score, we would just convert it to an undirected graph and calculate the score. Note that, unlike with degree centralization which has two scores (i.e. indegree centralization and outdegree centralization), there is only a single centralization score for a directed graph.

```{r}
d_graph
```

For the graph above, we can compare the centralization scores based on the different measures of centrality. The indegree centralization score is 0.428, the outdegree centralization score is 0.125, and the closeness centralization score is 0.556. What can we say about the closeness centralization scores for each type of centrality?

## Test your Knowledge

-   What does "closeness centrality" measure in a network?
-   How does closeness centrality differ from degree centrality when analyzing a network?
-   Explain why individuals with high closeness centrality might hold important roles within a network, even if they are not directly connected to many others.
-   Describe a scenario in which using closeness centrality would be more informative for a crime analyst than using degree centrality.
-   How is the calculation of closeness centrality different in directed graphs compared to undirected graphs?
-   How does the closeness centralization score help in determining the overall structure of a network?
-   What does it mean if the closeness centralization score of a graph is close to 1? What does it mean if the score is close to 0?

## Summary

In this chapter we focused on closeness centrality. This concept is particularly useful for understanding the efficiency of information flow in a network. Unlike degree centrality, which measures direct connections, closeness centrality provides insight into the influence and access a node has through indirect connections. Studies, such as the analysis of drug trafficking and terrorist networks, have shown that individuals with high closeness centrality, though not necessarily the most directly connected, play crucial roles in maintaining communication and operational efficiency. In the next chapter we will focus on a final conceptualization and operationalization of centrality: *betweenness*.
